---
title: "ToothGrowth Analysis"
author: "Harland Hendricks"
date: "June 27, 2018"
output:
  pdf_document: 
    fig_caption: yes
    fig_height: 4
    fig_width: 4
  html_document:
    df_print: paged
---

The instructions for this assignment are located [here.](https://www.coursera.org/learn/statistical-inference/peer/3k8j5/statistical-inference-course-project)

Built with R version `r getRversion()` with the following system:

```{r, echo = FALSE}
Sys.info()
```

Load the required libraries

```{r}
library(dplyr)
library(ggplot2)
library(gridExtra)
```

## Analysis of ToothGrowth supp / dose

### Load the Data

Load the data into object named "dat"

```{r}
data("ToothGrowth")
dat <- ToothGrowth
```

### Structure and summary of dat

*Note: Use ?ToothGrowth to find information on the dataset.* 

```{r}
str(dat)
summary(dat)
```

### Exploratory Data Analysis

```{r fig.cap = "Comparison of Orange Juice and Ascorbic Acid on Tooth Growth Overall and by Doses"}
supplen <- ggplot(dat, aes(x = supp, y = len)) + geom_boxplot()
doselen <- ggplot(dat, aes(x = interaction(supp, dose), y = len)) + geom_boxplot()
grid.arrange(supplen, doselen, ncol = 2)
```

Figure 1 shows that Orange Juice may have increased tooth growth overall.  The 1mg/day 
doseage shows a significant increase, which we will test.

### Test of overall effect of Orange Juice and Ascorbic Acid on Tooth Growth

We test the null hypothesis that VC and OJ have a difference in means of 0:

```{r}
##Group data by OJ and VC
datoj <- filter(dat, supp == "OJ")
datvc <- filter(dat, supp == "VC")
##Run t-test
t.test(datoj$len - datvc$len, mu = 0, paired = FALSE, alternative = "two.sided", 
       conf.level = 0.95)
```

The results of the test indicate that we must reject the null hyopothesis due to the 
significance of OJ on tooth growth compared to VC.

### Test 0.5 and 1.0 mg/day doseage to see which one is more significant.

Since we determined that OJ significantly affects tooth growth, we now want to 
test doseage levels.  The 2 mg/day doseage looks like there is no effect, so we will
test that first to confirm there is no significance:

```{r}
##Group data by doseage
datojd1 <- filter(datoj, dose == 0.5)
datojd2 <- filter(datoj, dose == 1)
datojd3 <- filter(datoj, dose == 2)
datvcd1 <- filter(datvc, dose == 0.5)
datvcd2 <- filter(datvc, dose == 1)
datvcd3 <- filter(datvc, dose == 2)
##Run t.test on doseage of 2 mg/day
t.test(datojd3$len - datvcd3$len, mu = 0, paired = FALSE, alternative = "two.sided", 
       conf.level = 0.95)
```

This test indicates that there is not enough evidence to reject the null hypothesis, 
therefore the difference in means is 0.

Test of doseage 0.5 mg/day:

```{r}
t.test(datojd1$len - datvcd1$len, mu = 0, paired = FALSE, alternative = "two.sided", 
       conf.level = 0.95)
```

Test of doseage 1 mg/day:

```{r}
t.test(datojd2$len - datvcd2$len, mu = 0, paired = FALSE, alternative = "two.sided", 
       conf.level = 0.95)
```

Both these test show that OJ significantly improved tooth growth.  The null hypothesis 
that the difference in means between OJ and VC is 0 should be rejected.

When comparing both, the doseage of 1 mg/day looks like a more effective treatment
due to the significance of the test.

### Conclussion and Assumptions

After identifying potential significant treatments for tooth decay during our exploratory 
analysis, we ran T-Tests on the data to confirm whether or not OJ is a better treatment than
Ascorbic Acid, and which doseage was more effective.

We can conclude that OJ does improve tooth growth in guinea pigs with a 95% significance
level.

We can also conclude that OJ in a 1 mg/day doseage is the most significant treatment doseage.  
Alternatively, we can say that increasing doseage to 2 mg/day reduces the effectiveness
of OJ on tooth growth.

The test performed assume that the variances were not equal and that the data was normally 
distributed.